---
title: "Using the Tesseract OCR engine in R"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Using the Tesseract OCR engine in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = "")
options(width = 100, max.print = 100)
has_nld <- "nld" %in% tesseract::tesseract_info()$available
```

The tesseract package provides R bindings [Tesseract](https://opensource.google.com/projects/tesseract): a powerful optical character recognition (OCR) engine that supports over 100 languages. The engine is highly configurable in order to tune the detection algorithms and obtain the best possible results.

Keep in mind that OCR (pattern recognition in general) is a very difficult problem for computers. Results will rarely be perfect and the accuracy rapidly decreases with the quality of the input image. But if you can get your input images to reasonable quality, Tesseract can often help to extract most of the text from the image.

## Extract Text from Images

OCR is the process of finding and recognizing text inside images, for example from a screenshot, scanned paper. The image below has some example text:


```{r echo=FALSE}
magick::image_read('http://jeroen.github.io/images/testocr.png')
```

```{r}
library(tesseract)
eng <- tesseract("eng")
text <- tesseract::ocr("http://jeroen.github.io/images/testocr.png", engine = eng)
cat(text)
```

Not bad! The `ocr_data()` function returns all words in the image along with a bounding box and confidence rate.

```{r}
results <- tesseract::ocr_data("http://jeroen.github.io/images/testocr.png", engine = eng)
print(results, n = 20)
```

## Language Data

The tesseract OCR engine uses language-specific training data in the recognize words. The OCR algorithms bias towards words and sentences that frequently appear together in a given language, just like the human brain does. Therefore the most accurate results will be obtained when using training data in the correct language. 

Use `tesseract_info()` to list the languages that you currently have installed.

```{r}
tesseract_info()
```

By default the R package only includes English training data. Windows and Mac users can install additional training data using `tesseract_download()`. Let's OCR a screenshot from Wikipedia in Dutch (Nederlands) 

```{r, eval=FALSE}
tesseract_download("nld", progress = FALSE)
```

```{r eval = has_nld}
dutch <- tesseract("nld")
text <- ocr("https://jeroen.github.io/images/utrecht.png", engine = dutch)
cat(text)
```

As you can see immediately: almost perfect! (OK just take my word). 

## Tesseract Parameters

Tesseract supports hundreds of [control parameters](https://github.com/tesseract-ocr/tesseract/wiki/ControlParams) which alter the OCR engine. Use `tesseract_params()` to list all parameters with their default value and a brief description. It also has a handy `filter` argument to quickly find parameters that match a particular string.

```{r}
# List all parameters with *colour* in name or description
tesseract_params('colour')
```

One powerful parameter is `tessedit_char_whitelist` which restricts the output to a limited set of characters. This may be useful for reading for example numbers such as a bank account, zip code, or gas meter.

```{r echo=FALSE}
magick::image_read('https://jeroen.github.io/images/receipt.png', strip = TRUE)
```


```{r}
numbers <- tesseract(options = list(tessedit_char_whitelist = "$.0123456789"))
cat(ocr("https://jeroen.github.io/images/receipt.png", engine = numbers))
```

To test if this actually works, look what happens if we remove the `$` from `tessedit_char_whitelist`:

```{r}
# Do not allow any dollar sign 
numbers2 <- tesseract(options = list(tessedit_char_whitelist = ".0123456789"))
cat(ocr("https://jeroen.github.io/images/receipt.png", engine = numbers2))
```

Notice how this forces tesseract to detect a number (3 or 8 or 5) if we rule out the dollar sign.

A config file is simply a text file parameters, one name and value per line. You can load config files from your current working directory, or load one of the standard tesseract config files from the tessdata directory.

## Read from PDF files

If your images are stored in PDF files they first need to be converted to a proper image format. We can do this in R using the `pdf_convert` function from the pdftools package. Use a high DPI to keep quality of the image.

```{r, eval=require(pdftools)}
pngfile <- pdftools::pdf_convert('https://jeroen.github.io/images/ocrscan.pdf', dpi = 600)
text <- tesseract::ocr(pngfile)
cat(text)
```

## Preprocessing with Magick

The accuracy of the OCR process depends on the quality of the input image. You can often improve results by properly scaling the image, removing noise and artifacts or cropping the area where the text it.


```{r}
library(magick)
(input <- image_read("https://jeroen.github.io/images/bowers.jpg"))

text <- input %>%
  image_resize("2000") %>%
  image_convert(colorspace = 'gray') %>%
  image_trim(fuzz = 10) %>%
  image_ocr()

cat(text)
```

